---
title: "How to calculate portfolio returns in R"
author: "DD"
date: "2018-04-05"
description: "Calculating portfolio returns in R"
output: html_document
---

## Calculating portfolio returns in R

In this post we will learn to calculate portfolio returns using R. Initially we will do this manually and then use the `tidyquant` package to calculate the portfolio returns for our purpose.

### Calculating portfolio returns using the formula

A portfolio return is the weighted average of indidual assets in the portfolio. 

Here is what we need

- Asset symbols that make up our portfolio
- Price data for the assets
- weights of assets
- Calculating the weighted average of our assets returns
- Adding them to get the portfolio returns


Lets first load the packages

```{r message=FALSE, warning=FALSE}

library(tidyquant)

```

We will invest in the following assets

- Aggregate Bonds ETF (BND) 10%
- Small Cap ETF (VB) 20%
- Developed markets ETF (VEA) 25%
- S&P 500 ETF (VOO)  25%
- Emerging Markets ETF (VWO) 20% 

So lets assign our ticker symbols and weights.

```{r}

# Asset tickers
tickers = c('BND', 'VB', 'VEA', 'VOO', 'VWO')

# Asset weights

wts = c(0.1,0.2,0.25,0.25,0.2)

```

Next we will download the price data

```{r}

price_data <- tq_get(tickers,
                     from = '2013-01-01',
                     to = '2018-03-01',
                     get = 'stock.prices')

```

Then we will calculate the returns for our assets. We will calculate the daily returns.

```{r}

ret_data <- price_data %>%
  group_by(symbol) %>%
  tq_transmute(select = adjusted,
               mutate_fun = periodReturn,
               period = "daily",
               col_rename = "ret")
  

```

Lets take a look at the first and second row of each asset in our returns data.

```{r}

ret_data %>%
  group_by(symbol) %>%
  slice(c(1,2))

```

Next we will have to add the weight column to our returns data. For that we will create a new weights table and then join that to our returns data table.

```{r}

wts_tbl <- tibble(symbol = tickers,
                  wts = wts)
head(wts_tbl)

  

```
Now lets join this to our returns data table.

```{r}

ret_data <- left_join(ret_data,wts_tbl, by = 'symbol')

```

Again lets look at the first and second row of our table.

```{r}

ret_data %>%
  group_by(symbol) %>%
  slice(c(1,2))

```

We can see that the weights have been correctly added to the corresponding assets. Now lets mutlply the two columns to get our weighted average.

```{r}

ret_data <- ret_data %>%
  mutate(wt_return = wts * ret)

```

Lets take a quick look at the result of our operation.

```{r}

ret_data %>%
  group_by(symbol) %>%
  slice(c(1,2))

```

We now have a weighted returns column. Now the portfolio return for each day is simply the sum of the weighted returns each day. Lets add that and get our portfolio returns.

```{r}

port_ret <- ret_data %>%
  group_by(date) %>%
  summarise(port_ret = sum(wt_return))

head(port_ret)

```

Now we have just calculated the portfolio returns using a manual process. Lets visualize the returns.

```{r}

port_ret %>%
  ggplot(aes(x = port_ret)) + 
  geom_histogram(bins = 60) +
  theme_light() +
  labs(x = "Portfolio Returns",
       y = "Frequency",
       title = "Daily Portfolio returns histogram")

```

We will analyze the portfolio returns little more in the post but right now we will show you a simpler way to calculate the portfolio returns using the `tidyquant` package.

`tidyquant` helps us eliminate the extra steps we took to add the weights columns and do the multiplication and the additions.

So lets see how we will do this in `tidyquant`

```{r}

port_ret_tidy <- price_data %>%
  group_by(symbol) %>%
  tq_transmute(select = adjusted,
               mutate_fun = periodReturn,
               period = "daily",
               col_rename = "ret") %>%
  #Using tq_portfolio from tidyquant
  tq_portfolio(assets_col = symbol,
               returns_col = ret,
               weights = wts,
               geometric = FALSE,
               col_rename = 'port_ret')
  


```


As you will notice the first part is similar to what we did before, but then we simply used `tq_portfolio` to calculate our portfolio returns. Here we have to specify the asset column, the returns column and the weights to calculate the portfolio returns. We will demonstrate that the returns are the same by taking the difference between two methods. First we show the histogram of the returns and the we plot the time series.

```{r}

port_ret_tidy %>%
  ggplot(aes(x = port_ret)) + 
  geom_histogram(bins = 60) +
  theme_light() +
  labs(x = "Portfolio Returns",
       y = "Frequency",
       title = "Daily Portfolio returns histogram\nCalculated using tidyquant")
  

```


```{r}

diff = port_ret$port_ret - port_ret_tidy$port_ret

unique(diff)

```
We can see the difference between the two methods is 0.

### Portfolio mean returns and standard deviation

Now that we are confident about our work, lets calculate the mean returns and the standard deviation of the portfolio.

```{r}

mean(port_ret_tidy$port_ret, na.rm = TRUE)

```


```{r}

sd(port_ret_tidy$port_ret, na.rm = TRUE)

```


We will conclude this post here and in the next post we will analyze our portfolio futher.

## Summary

In this post we learned

- To download asset prices from yahoo finance
- To calculate portfolio returns manually
- To calculate portfolio returns using `tidyquant`
- To calculate portfolio mean returns and portfolio standard deviation


