---
title: "Quantitative Investment Analysis - Chapter 6"
author: "DD"
date: "2020-05-02"
description: "Chapter 6: Sampling and Estimation"
output:
  html_document:
    highlight: tango
---

```{r message=FALSE, warning=FALSE}

library(tidyverse)
library(knitr)
library(kableExtra)

```

In this post we will solve the end of the chapter practice problems from chapter 6 of the [book](https://www.amazon.com/Quantitative-Investment-Analysis-CFA-Institute/dp/111910422X).

### Problem 1

Peter Biggs wants to know how growth managers performed last year. Biggs
assumes that the population cross-sectional
standard deviation of growth manager
returns is 6 percent and that the returns are independent across managers.

- A How large a random sample does Biggs need if he wants the standard deviation
of the sample means to be 1 percent?
- B How large a random sample does Biggs need if he wants the standard deviation
of the sample means to be 0.25 percent?

We know that

![](https://www.gstatic.com/education/formulas/images_long_sheet/en/standard_error.svg)

```{r}

# Population sd is 0.06
pop_sd <- 6/100

# The standard deviation or standard error of the sample mean is σX = σ / sqrt(n)
# 0.01 = pop_sd / sqrt(n)

n = (pop_sd / 0.01) ** 2

cat("Biggs need", n, "samples, if he wants the standard deviation
    of the sample means to be 1 percent")

```

A - Biggs need 36 samples, if he wants the standard deviation of the sample means to be 1 percent

```{r}

n = (pop_sd / 0.0025) ** 2

cat("Biggs need", n, "samples, if he wants the standard deviation
    of the sample means to be 0.25 percent")

```

B - Biggs need 576 samples, if he wants the standard deviation of the sample means to be 0.25 percent

As standard error decreases, the sample size increases.

### Problem 2

Petra Munzi wants to know how value managers performed last year. Munzi
estimates that the population cross-sectional
standard deviation of value manager
returns is 4 percent and assumes that the returns are independent across
managers.

- A Munzi wants to build a 95 percent confidence interval for the mean return.
How large a random sample does Munzi need if she wants the 95 percent
confidence interval to have a total width of 1 percent?
- B Munzi expects a cost of about $10 to collect each observation. If she has
a $1,000 budget, will she be able to construct the confidence interval she
wants?
