---
title: "How to download stock prices in R"
author: "DD"
date: "2018-04-02"
description: "Downloading stock prices in R"
output: html_document
---

## Getting stock prices from Yahoo Finance

One of the most important tasks in financial markets is to analyze historical returns on various investments. To perform this analysis we need historical data for the assets. There are many data providers, some are free most are paid. In this chapter we will use the data from Yahoo's finance website. Since Yahoo was bought by Verizon, there have been several changes with their API. They may decide to stop providing stock prices in the future. So the method discussed on this article may not work in the future.


## R packages to download stock price data

There are several ways to get financial data into R. The most popular method is the `quantmod` package. You can install it by typing the command `install.packages("quantmod")` in your R console. The prices downloaded in by using `quantmod` are `xts` `zoo` objects. For our calculations we will use `tidyquant` package which downloads prices in a tidy format as a `tibble`. You can download the `tidyquant` package by typing `install.packages("tidyquant")` in you R console. `tidyquant` includes `quantmod` so you can install just `tidyquant` and get the `quantmod` packages as well.

Lets load the library first.

```{r message=FALSE,warning=FALSE}

library(tidyquant)

```

First we will download Apple price using `quantmod` from January 2017 to February 2018. By default `quantmod` download and stores the symbols with their own names. You can change this by passing the argument `auto.assign = FALSE`.

```{r}

options("getSymbols.warning4.0"=FALSE)
options("getSymbols.yahoo.warning"=FALSE)
# Downloading Apple price using quantmod

getSymbols("AAPL", from = '2017-01-01',
           to = "2018-03-01",warnings = FALSE,
           auto.assign = TRUE)


```

Lets look at the first few rows.

```{r}

head(AAPL)

```

Lets look at the class of this object.

```{r}

class(AAPL)



```

As we mentioned before this is an `xts` `zoo` object. We can also chart the Apple stock price. We just pass the command `chart_Series`

```{r}

chart_Series(AAPL)

```

We can even zoom into a certain period of the series. Lets zoom in on the Dec to Feb period.

```{r}

chart_Series(AAPL['2017-12/2018-03'])

```

We can download prices for several stocks. There are several steps to this

```{r}

tickers = c("AAPL", "NFLX", "AMZN", "K", "O")

getSymbols(tickers,
           from = "2017-01-01",
           to = "2017-01-15")

prices <- map(tickers,function(x) Ad(get(x)))
prices <- reduce(prices,merge)
colnames(prices) <- tickers

```


```{r}

head(prices)
class(prices)

```

But we prefer the `tidyquant` package to download stock prices. Below we will demonstrate the simplicity of the process.

```{r}

aapl <- tq_get('AAPL',
               from = "2017-01-01",
               to = "2018-03-01",
               get = "stock.prices")


```

```{r}

head(aapl)

```


```{r}

class(aapl)

```

We can see that the object `aapl` is a `tibble`. Next we can chart the price for Apple. For that we will use the very popular `ggplot2` package.

```{r}

aapl %>%
  ggplot(aes(x = date, y = adjusted)) +
  geom_line() +
  theme_classic() +
  labs(x = 'Date',
       y = "Adjusted Price",
       title = "Apple price chart") +
  scale_y_continuous(breaks = seq(0,300,10))

```

We can also download multiple stock prices.

```{r}

tickers = c("AAPL", "NFLX", "AMZN", "K", "O")

prices <- tq_get(tickers,
                 from = "2017-01-01",
                 to = "2017-03-01",
                 get = "stock.prices")

```


```{r}

head(prices)

```


This data is in tidy format, where symbols are stacked on top of one another. To see the first row of each symbol, we need to slice the data.

```{r}

prices %>%
  group_by(symbol) %>%
  slice(1)

```

We can also chart the time series of all the prices.

```{r}

prices %>%
  ggplot(aes(x = date, y = adjusted, color = symbol)) +
  geom_line()

```

This chart look weird, since the scale is not appropriate. Amazon price is above \$800, other stocks are under \$200. We can fix this with `facet_wrap`

```{r}

prices %>%
  ggplot(aes(x = date, y = adjusted, color = symbol)) +
  geom_line() +
  facet_wrap(~symbol,scales = 'free_y') +
  theme_classic() +
  labs(x = 'Date',
       y = "Adjusted Price",
       title = "Price Chart") +
  scale_x_date(date_breaks = "month",
               date_labels = "%b\n%y")



```

