---
title: "How to calculate Cumulative portfolio returns in R"
author: "DD"
date: "2018-04-10"
description: "Calculating cumulative portfolio returns in R"
output: html_document
---

## Calculating Cumulative portfolio returns in R

In the last post we learned how to construct a portfolio in R. We also learned how to calculate the daily portfolio returns. In this post we will learn how to calculate portfolio cumulative returns.

First lets load the library.

```{r message=FALSE, warning=FALSE}

library(tidyquant)

```

Then lets load the ticker symbols for our assets that we will include in our portfolio.

```{r}
# Asset tickers

tickers = c('BND', 'VB', 'VEA', 'VOO', 'VWO')

```

We will also create a vector for our asset weights.

```{r}
# Asset weights

wts = c(0.1,0.2,0.25,0.25,0.2)


```

Next lets download the price data from yahoo finance.

```{r}

price_data <- tq_get(tickers,
                     from = '2013-01-01',
                     to = '2018-03-01',
                     get = 'stock.prices')

```

Next we will calculate the daily returns for our assets.

```{r}

ret_data <- price_data %>%
  group_by(symbol) %>%
  tq_transmute(select = adjusted,
               mutate_fun = periodReturn,
               period = "daily",
               col_rename = "ret")

```

For our ease of calculations, we will create a weight table.

```{r}

wts_tbl <- tibble(symbol = tickers,
                  wts = wts)

```

Then we will join our weights table and the returns data.

```{r}

ret_data <- left_join(ret_data,wts_tbl, by = 'symbol')

```

We will then calculate the weighted average of our asset returns.

```{r}

ret_data <- ret_data %>%
  mutate(wt_return = wts * ret)

```

Finally the portfolio returns are the sum of the weighted returns.

```{r}

port_ret <- ret_data %>%
  group_by(date) %>%
  summarise(port_ret = sum(wt_return))

```

Once we have the portfolio returns, we will use the `cumprod()` function to calculate the cumulative returns for the portfolio.

```{r}

port_cumulative_ret <- port_ret %>%
  mutate(cr = cumprod(1 + port_ret))

```

We can visualize the portfolio returns using the below code.

```{r}

port_cumulative_ret %>%
  ggplot(aes(x = date, y = cr)) +
  geom_line() +
  labs(x = 'Date',
       y = 'Cumulative Returns',
       title = 'Portfolio Cumulative Returns') +
  theme_classic() +
  scale_y_continuous(breaks = seq(1,2,0.1)) +
  scale_x_date(date_breaks = 'year',
               date_labels = '%Y')

```

We will post the entire code here.

```{r eval=FALSE}

library(tidyquant)

# Asset tickers

tickers = c('BND', 'VB', 'VEA', 'VOO', 'VWO')

# Asset weights

wts = c(0.1,0.2,0.25,0.25,0.2)

price_data <- tq_get(tickers,
                     from = '2013-01-01',
                     to = '2018-03-01',
                     get = 'stock.prices')

ret_data <- price_data %>%
  group_by(symbol) %>%
  tq_transmute(select = adjusted,
               mutate_fun = periodReturn,
               period = "daily",
               col_rename = "ret")

wts_tbl <- tibble(symbol = tickers,
                  wts = wts)
ret_data <- left_join(ret_data,wts_tbl, by = 'symbol')

ret_data <- ret_data %>%
  mutate(wt_return = wts * ret)

port_ret <- ret_data %>%
  group_by(date) %>%
  summarise(port_ret = sum(wt_return))

port_cumulative_ret <- port_ret %>%
  mutate(cr = cumprod(1 + port_ret))

```

But there is a simpler code if we use the tidyquant function. We will demonstrate how to use to tidyquant to calculate cumulative portfolio returns.

```{r}

port_ret_tidyquant <- ret_data %>%
  tq_portfolio(assets_col = symbol,
               returns_col = ret,
               weights = wts,
               col_rename = 'port_ret',
               geometric = FALSE)

port_cumulative_ret_tidyquant <- port_ret_tidyquant %>%
  mutate(cr = cumprod(1 + port_ret))

```

As we can see that the above code is much shorter than the previous code. We can compare the two portfolio calculations.

```{r}

port_cumulative_ret %>%
  mutate(port_ret_tidyquant = port_cumulative_ret_tidyquant$cr) %>%
  select(-port_ret) %>%
  rename(long_method = cr) %>%
  gather(long_method,port_ret_tidyquant,
         key = port_method,
         value = cr) %>%
  ggplot(aes(x = date, y = cr, color = port_method)) +
  geom_line() +
  labs(x = 'Date',
       y = 'Cumulative Returns',
       title = 'Portfolio Cumulative Returns') +
  theme_classic() +
  scale_y_continuous(breaks = seq(1,2,0.1)) +
  scale_x_date(date_breaks = 'year',
               date_labels = '%Y')

```

We can see that both line overlap each other and the returns are the same. So in the future its best to use the shorter `tq_portfolio()` method.
